---
title: "List of useful epidemiology R packages"
output: 
  html_notebook: 
    toc: yes
---

Below is a long list of suggested packages for common epidemiological work in R. You can copy this code, run it, and all of these packages will install from CRAN and load for use in the current R session. If a package is already installed, it will be loaded for use only.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
##########################################
# List of useful epidemiology R packages #
##########################################

# This script uses the p_load() function from pacman R package, 
# which installs if package is absent, and loads for use if already installed


# Ensures the package "pacman" is installed
if (!require("pacman")) install.packages("pacman")


# Packages available from CRAN
##############################
pacman::p_load(
     
     # learning R
     ############
     learnr,   # interactive tutorials in RStudio Tutorial pane
     swirl,    # interactive tutorials in R console
        
     # project and file management
     #############################
     here,     # file paths relative to R project root folder
     rio,      # import/export of many types of data
     openxlsx, # import/export of multi-sheet Excel workbooks 
     
     # package install and management
     ################################
     pacman,   # package install/load
     renv,     # managing versions of packages when working in collaborative groups
     remotes,  # install from github
     
     # General data management
     #########################
     tidyverse,    # includes many packages for tidy data wrangling and presentation
          #dplyr,      # data management
          #tidyr,      # data management
          #ggplot2,    # data visualization
          #stringr,    # work with strings and characters
          #forcats,    # work with factors 
          #lubridate,  # work with dates
          #purrr       # iteration and working with lists
     linelist,     # cleaning linelists
     naniar,       # assessing missing data
     
     # statistics  
     ############
     janitor,      # tables and data cleaning
     gtsummary,    # making descriptive and statistical tables
     rstatix,      # quickly run statistical tests and summaries
     broom,        # tidy up results from regressions
     lmtest,       # likelihood-ratio tests
     easystats,
          # parameters, # alternative to tidy up results from regressions
          # see,        # alternative to visualise forest plots 
     
     # epidemic modeling
     ###################
     epicontacts,  # Analysing transmission networks
     EpiNow2,      # Rt estimation
     EpiEstim,     # Rt estimation
     projections,  # Incidence projections
     incidence2,   # Make epicurves and handle incidence data
     i2extras,     # Extra functions for the incidence2 package
     epitrix,      # Useful epi functions
     distcrete,    # Discrete delay distributions
     
     
     # plots - general
     #################
     #ggplot2,         # included in tidyverse
     cowplot,          # combining plots  
     # patchwork,      # combining plots (alternative)     
     RColorBrewer,     # color scales
     ggnewscale,       # to add additional layers of color schemes

     
     # plots - specific types
     ########################
     DiagrammeR,       # diagrams using DOT language
     incidence2,       # epidemic curves
     gghighlight,      # highlight a subset
     ggrepel,          # smart labels
     plotly,           # interactive graphics
     gganimate,        # animated graphics 

     
     # gis
     ######
     sf,               # to manage spatial data using a Simple Feature format
     tmap,             # to produce simple maps, works for both interactive and static maps
     OpenStreetMap,    # to add OSM basemap in ggplot map
     spdep,            # spatial statistics 
     
     # routine reports
     #################
     rmarkdown,        # produce PDFs, Word Documents, Powerpoints, and HTML files
     reportfactory,    # auto-organization of R Markdown outputs
     officer,          # powerpoints
     
     # dashboards
     ############
     flexdashboard,    # convert an R Markdown script into a dashboard
     shiny,            # interactive web apps
     
     # tables for presentation
     #########################
     knitr,            # R Markdown report generation and html tables
     flextable,        # HTML tables
     #DT,              # HTML tables (alternative)
     #gt,              # HTML tables (alternative)
     #huxtable,        # HTML tables (alternative) 
     
     # phylogenetics
     ###############
     ggtree,           # visualization and annotation of trees
     ape,              # analysis of phylogenetics and evolution
     treeio            # to visualize phylogenetic files
 
)
```

```{r}
# Packages to download from Github (not available on CRAN)
##########################################################

# Development version of epicontacts (for transmission chains with a time x-axis)
pacman::p_install_gh("reconhub/epicontacts@timeline")

# The package for this handbook, which includes all the example data  
pacman::p_install_gh("appliedepi/epirhandbook")
```


```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("treeio")
BiocManager::install("ggtree")
```

```{r}
#install easystats
install.packages(c("insight", "datawizard", "bayestestR", "performance",
                    "parameters", "effectsize", "correlation", "modelbased",
                    "see", "report"), repos = "https://easystats.r-universe.dev")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

